---
title: "General elections 9 April 2006 (Chamber of Deputy and Senate of the Republic)"
author: "Francesco Bailo"
date: "29 January 2019"
output: 
  rmarkdown::github_document
---

The details on variables and decriptive statistics on the results of the 2006 elections for both the Chamber and the Senate (formally two different elections with different electorates).

# Count datasets

Count data for geographic entitiy (municipal level or lower) is provided in the following files:

* `open_csv_20060409_senato_geo_count.csv`;
* `open_csv_20060409_camera_geo_count.csv`;
* `open_csv_20060409_senato_party_count.csv`;
* `open_csv_20060409_camera_party_count.csv`.


```{r, include=F}
setwd("~/public_git/ScrapeOpen/elezionistorico.interno.gov.it")
camera_party_count <- read.csv('open_csv_20060409_camera_party_count.csv')
camera_geo_count <- read.csv('open_csv_20060409_camera_geo_count.csv')

senato_party_count <- read.csv('open_csv_20060409_senato_party_count.csv')
senato_geo_count <- read.csv('open_csv_20060409_senato_geo_count.csv')
```

For each election, Chamber of Deputies (`camera`) and the Senate of the Republic (`senato`), 2 open datasets are provided: 

* `party_count` with information on votes received by parties.
* `geo_count` with general electoral count information on the geographic entity.  

## Variable description

### Character variables
* `election` either `Senato` or `Camera`.
* `date` in format `%Y-%m-%d`.
* `geo_lev_1` to `geo_lev_2` the label as set by the original data provider for the the electoral boundaries (from larger to smaller) of the geographical entity described by the observation.
* `geo_entity` name of the geographical entity described by the observation as set by the original data provider.

### Numerical variables
* `party_votes` the sum of votes received by a party in the geographical entity.
* `candidate_votes` the sum of votes received by a candidate (that is, also by a coalition) in the geographical entity.
* `registered_voters` the sum of voters registered for the election and entitled to vote in the geographical entity.
* `blank_votes`
* `invalid_votes` the sum ofnon-assigned votes (it includes `blank_votes`) in the geographical entity.

## Variable summary

### camera_party_count

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  camera_party_count %>%
  summarize(party_votes = sum(party_votes),
            `%` = 100)
kable(tb, format.args = list(big.mark = ','), digits=2)
```

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  camera_party_count %>%
  group_by(party) %>%
  summarize(party_votes = sum(party_votes),
            `%` = sum(party_votes) / sum(camera_party_count$party_votes) * 100) %>%
  arrange(-party_votes)
kable(tb, format.args = list(big.mark = ','), digits=2)
```

### senato_party_count

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  senato_party_count %>%
  summarize(party_votes = sum(party_votes),
            `%` = 100)
kable(tb, format.args = list(big.mark = ','), digits=2)
```

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  senato_party_count %>%
  group_by(party) %>%
  summarize(party_votes = sum(party_votes),
            `%` = sum(party_votes) / sum(senato_party_count$party_votes) * 100) %>%
  arrange(-party_votes)
kable(tb, format.args = list(big.mark = ','), digits=2)
```

### camera_geo_count

```{r, echo=FALSE,warning=FALSE,message=FALSE}
tb <-
  camera_geo_count %>%
  summarize(registered_voters = sum(registered_voters),
            effective_voters = sum(effective_voters),
            `% (effective)` = sum(camera_geo_count$effective_voters) / sum(camera_geo_count$registered_voters) * 100,
            blank_votes = sum(blank_votes),
            invalid_votes = sum(invalid_votes), 
            `% (invalid)` = sum(camera_geo_count$invalid_votes) / sum(camera_geo_count$effective_voters) * 100)
kable(tb, format.args = list(big.mark = ','), digits=2)

```

### senato_geo_count

```{r, echo=FALSE,warning=FALSE,message=FALSE}
tb <-
  senato_geo_count %>%
  summarize(registered_voters = sum(registered_voters),
            effective_voters = sum(effective_voters),
            `% (effective)` = sum(senato_geo_count$effective_voters) / sum(senato_geo_count$registered_voters) * 100,
            blank_votes = sum(blank_votes),
            invalid_votes = sum(invalid_votes), 
            `% (invalid)` = sum(senato_geo_count$invalid_votes) / sum(senato_geo_count$effective_voters) * 100)
kable(tb, format.args = list(big.mark = ','), digits=2)

```


