---
title: "General elections 4 March 2018 (Chamber of Deputy and Senate of the Republic)"
author: "Francesco Bailo"
date: "3 June 2018"
output: 
  rmarkdown::github_document
---

The details on variables and decriptive statistics on the results of the 2018 elections. For both the Chamber and the Senate (formally two different elections with different electorates), voters had the option of either voting a *party* or the coaltion's *candidate* either in the case of monocoalitions (i.e. one-party coalition). Consequentially, for each coalition the votes received by a candidate must equal or be greater than the sum of votes received by the parties in the coalition. 

# Count datasets

Count data for geographic entitiy (municipal level or lower) is provided in the following files:

* `open_csv_20180304_camera_party_count.csv`;
* `open_csv_20180304_camera_candidate_count.csv`;
* `open_csv_20180304_camera_geo_count.csv`;
* `open_csv_20180304_senato_party_count.csv`;
* `open_csv_20180304_senato_candidate_count.csv`;
* `open_csv_20180304_senato_geo_count.csv`;

```{r, include=F}
camera_party_count <- read.csv('open_csv_20180304_camera_party_count.csv')
camera_candidate_count <- read.csv('open_csv_20180304_camera_candidate_count.csv')
camera_geo_count <- read.csv('open_csv_20180304_camera_geo_count.csv')

senato_party_count <- read.csv('open_csv_20180304_senato_party_count.csv')
senato_candidate_count <- read.csv('open_csv_20180304_senato_candidate_count.csv')
senato_geo_count <- read.csv('open_csv_20180304_senato_geo_count.csv')
```

For each election, Chamber of Deputies (`camera`) and the Senate of the Republic (`senato`), 3 open datasets are provided: 

* `party_count` with information on votes received by parties.
* `candidate_count` with information on votes received by candidates (that is, coalitions since each candidate is the uninominal candidate of a coaltion).
* `geo_count` with general electoral count information on the geographic entity.  

## Variable description

### Character variables
* `election` either `Senato` or `Camera`.
* `date` in format `%Y-%m-%d`.
* `geo_lev_1` to `geo_lev_4` the label as set by the original data provider for the the electoral boundaries (from larger to smaller) of the geographical entity described by the observation.
* `geo_entity` name of the geographical entity described by the observation as set by the original data provider.
* `candidate` name of the candidate as set by the original data provider.
* `coalition` name of the coalition as set by the original data provider (in the case of a monocoalition, is the name of the party).
* `istat_nam` name of `geo_entity` as set by [Italy's National Institute of Statistics (Istat)](https://en.wikipedia.org/wiki/National_Institute_of_Statistics_(Italy)). 
* `minint_cod` identifier of the `geo_entity` as set by the Ministry of Interior (use this to join with the geographic files `open_shp_20180304_camera` and `open_shp_20180304_senato`). 

### Numerical variables
* `party_votes` the sum of votes received by a party in the geographical entity.
* `candidate_votes` the sum of votes received by a candidate (that is, also by a coalition) in the geographical entity.
* `registered_voters` the sum of voters registered for the election and entitled to vote in the geographical entity.
* `blank_votes`
* `invalid_votes` the sum ofnon-assigned votes (it includes `blank_votes`) in the geographical entity.

### Variables linking to other resources
* `party_wd_id` Wikidata URI of the party, in the format `Q[0-9]+` (to visit the resource page append `http://www.wikidata.org/wiki/`).
* `coalition_wd_id`  Wikidata URI of the coalition, in the format `Q[0-9]+` (in case of monocoalition, it corresponds to `party_wd_id`).

## Variable summary

### camera_party_count

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  camera_party_count %>%
  summarize(party_votes = sum(party_votes),
            `%` = 100)
kable(tb, format.args = list(big.mark = ','), digits=2)
```

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  camera_party_count %>%
  group_by(party) %>%
  summarize(party_votes = sum(party_votes),
            `%` = sum(party_votes) / sum(camera_party_count$party_votes) * 100) %>%
  arrange(-party_votes)
kable(tb, format.args = list(big.mark = ','), digits=2)
```

### senato_party_count

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  senato_party_count %>%
  summarize(party_votes = sum(party_votes),
            `%` = 100)
kable(tb, format.args = list(big.mark = ','), digits=2)
```

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  senato_party_count %>%
  group_by(party) %>%
  summarize(party_votes = sum(party_votes),
            `%` = sum(party_votes) / sum(senato_party_count$party_votes) * 100) %>%
  arrange(-party_votes)
kable(tb, format.args = list(big.mark = ','), digits=2)
```


### camera_candidate_count

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  camera_candidate_count %>%
  summarize(candidate_votes = sum(candidate_votes),
            `%` =  100)
kable(tb, format.args = list(big.mark = ','), digits=2)
```

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  camera_candidate_count %>%
  group_by(coalition) %>%
  summarize(coalition_votes = sum(candidate_votes),
            `%` = sum(candidate_votes) / sum(camera_candidate_count$candidate_votes) * 100) %>%
  arrange(-coalition_votes)
kable(tb, format.args = list(big.mark = ','), digits=2)
```

### senato_candidate_count

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  senato_candidate_count %>%
  summarize(candidate_votes = sum(candidate_votes),
            `%` =  100)
kable(tb, format.args = list(big.mark = ','), digits=2)
```

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  senato_candidate_count %>%
  group_by(coalition) %>%
  summarize(coalition_votes = sum(candidate_votes),
            `%` = sum(candidate_votes) / sum(senato_candidate_count$candidate_votes) * 100) %>%
  arrange(-coalition_votes)
kable(tb, format.args = list(big.mark = ','), digits=2)
```

### camera_geo_count

```{r, echo=FALSE,warning=FALSE,message=FALSE}
tb <-
  camera_geo_count %>%
  summarize(registered_voters = sum(registered_voters),
            effective_voters = sum(effective_voters),
            `% (effective)` = sum(camera_geo_count$effective_voters) / sum(camera_geo_count$registered_voters) * 100,
            blank_votes = sum(blank_votes),
            invalid_votes = sum(invalid_votes), 
            `% (invalid)` = sum(camera_geo_count$invalid_votes) / sum(camera_geo_count$effective_voters) * 100)
kable(tb, format.args = list(big.mark = ','), digits=2)

```

### senato_geo_count

```{r, echo=FALSE,warning=FALSE,message=FALSE}
tb <-
  senato_geo_count %>%
  summarize(registered_voters = sum(registered_voters),
            effective_voters = sum(effective_voters),
            `% (effective)` = sum(senato_geo_count$effective_voters) / sum(senato_geo_count$registered_voters) * 100,
            blank_votes = sum(blank_votes),
            invalid_votes = sum(invalid_votes), 
            `% (invalid)` = sum(senato_geo_count$invalid_votes) / sum(senato_geo_count$effective_voters) * 100)
kable(tb, format.args = list(big.mark = ','), digits=2)

```


