---
title: "[Italian constitutional referendum 4 December 2016"
author: "Francesco Bailo"
date: "24 October 2018"
output: 
  rmarkdown::github_document
---

The details on variables and decriptive statistics on the results of the 2016 constitutional referendum at the comune level.

# Dataset

* `open_csv_20161204_F.csv`

```{r, include=F}
setwd('ScrapeOpen/elezionistorico.interno.gov.it')
open_csv_20161204_F <- read.csv('open_csv_20161204_F.csv')
```

## Variable description

### Character variables
* `type` is `SCRUTINI`.
* `election` is `Referendum`.
* `date` in format `%Y-%m-%d`.
* `geo_lev_1` is `ITALIA` for all records.
* `geo_lev_2` is the region name as set by the original data provider.
* `geo_lev_3` is the province name as set by the original data provider (there are 106 provinces).
* `geo_entity` name of the geographical entity described by the observation as set by the original data provider.

### Numerical variables
* `yes_votes` the sum of "yes" votes.
* `no_votes` the sum of votes "no" votes.
* `registered_voters` the sum of voters registered for the election and entitled to vote in the geographical entity.
* `effective_voters` the sum of voters actually showing up to vote in the geographical entity.
* `blank_votes`
* `invalid_votes` the sum of non-assigned votes (it includes `blank_votes`) in the geographical entity.

### Variables linking to other resources
* `istat_cod` is the name used by ISTAT in the geographic file `open_shp_20161204_F`.

## Variable summary

### open_csv_20161204_F

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)

tb <- 
  with(open_csv_20161204_F,
       data.frame(yes_votes = sum(yes_votes) ,
                  no_votes = sum(no_votes),
                  registered_voters = sum(registered_voters),
                  effective_voters = sum(effective_voters),
                  blank_votes = sum(blank_votes),
                  invalid_votes = sum(invalid_votes)))
kable(tb, format.args = list(big.mark = ','), digits=2)

tb <- 
with(open_csv_20161204_F,
  data.frame(yes_votes = (sum(yes_votes) / (sum(yes_votes) + sum(no_votes)))*100,
            no_votes = (sum(no_votes) / (sum(yes_votes) + sum(no_votes))) *100,
            registered_voters = NA,
            effective_voters = (sum(effective_voters) / sum(registered_voters))*100,
            blank_votes = (sum(blank_votes) / sum(effective_voters))*100,
            invalid_votes = (sum(invalid_votes) / sum(effective_voters))*100))
kable(tb, format.args = list(big.mark = ','), digits=2)

```

