---
title: "General elections 13 April 2008 (Chamber of Deputy)"
author: "Francesco Bailo"
date: "16 September 2021"
output: 
  rmarkdown::github_document
---

read_raw_csv_20080413.R

The details on variables and decriptive statistics on the results of the 2008 elections for the Chamber of Deputies only.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Count datasets

Count data for geographic entitiy (municipal level) is provided in the following files:

* `open_csv_20080413_camera_geo_count.csv`;
* `open_csv_20080413_camera_party_count.csv`.

```{r, include=F}
camera_geo_count <- read.csv('~/open_csv_20080413_camera_geo_count.csv')
camera_party_count <- read.csv('~/open_csv_20080413_camera_party_count.csv')
```

For the Chamber of Deputies (`camera`), 2 open datasets are provided: 

* `party_count` with information on votes received by parties ("liste" in Italian).
* `geo_count` with general electoral count information on the geographic entity.

## Variable description

### Character variables
* `election` set to `Camera`.
* `date` in format `%Y-%m-%d`.
* `geo_lev_1` to `geo_lev_2` the label as set by the original data provider for the the electoral boundaries (from larger to smaller) of the geographical entity described by the observation.
* `geo_entity` name of the geographical entity described by the observation as set by the original data provider.
* `party` name of the list as set by the original data provider.
* `istat_cod` code of `geo_entity` as set by [Italy's National Institute of Statistics (Istat)](https://en.wikipedia.org/wiki/National_Institute_of_Statistics_(Italy)) (also `PRO_COM_T`). Use this to join with the geographic files `raw_shp_20080413`.) 

### Numerical variables
* `party_votes` the sum of votes received by a party in the geographical entity.
* `tot_party_votes` the sum of votes received by all parties in the geographical entity.
* `registered_voters` the sum of voters registered for the election and entitled to vote in the geographical entity.
* `registered_male_voters`
* `effective_voters` the sum of voters actually casting a vote in the election
* `effective_male_voters`
* `blank_votes`
* `invalid_votes` the sum ofnon-assigned votes (it includes `blank_votes`) in the geographical entity.

## Variable summary

### camera_party_count

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  camera_party_count %>%
  summarize(party_votes = sum(party_votes),
            `%` = 100)
kable(tb, format.args = list(big.mark = ','), digits=2)
```

```{r, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
tb <- 
  camera_party_count %>%
  group_by(party) %>%
  summarize(party_votes = sum(party_votes),
            `%` = sum(party_votes) / sum(camera_party_count$party_votes) * 100) %>%
  arrange(-party_votes)
kable(tb, format.args = list(big.mark = ','), digits=2)
```

### camera_geo_count

```{r, echo=FALSE,warning=FALSE,message=FALSE}
tb <-
  camera_geo_count %>%
  summarize(registered_voters = sum(registered_voters),
            registered_male_voters = sum(registered_male_voters),
            effective_voters = sum(effective_voters),
                        `% (effective)` = sum(camera_geo_count$effective_voters) / sum(camera_geo_count$registered_voters) * 100,
            effective_male_voters = sum(effective_male_voters),
            tot_party_votes = sum(tot_party_votes),
            blank_votes = sum(blank_votes),
            `% (valid votes)` = sum(tot_party_votes) / sum(effective_voters) * 100)
kable(tb, format.args = list(big.mark = ','), digits=2)

```
